From big data to deep insight in developmental science

### 2015-07-30 wiley-big-data-devel/big-data.tex
thru govt run section

\subsection{Government-Supervised and Managed Datasets}

Some of the largest and most widely used datasets in developmental science are population-based, often with an epidemiological dimension.
Several notable examples began data collection in the 1950's or 60's.
The National Health and Nutrition Examination Survey (NHANES) \cite{NHANES} began collecting data in 1959 with a focus on characterizing the health and nutritional status of U.S. infants, children, and adults. 
The annual survey combines interviews about demographic, socioeconomic, dietary and health related topics and physical examinations involving medical, dental, and physiological measurements, as well as laboratory tests.
The study assesses a nationally representative sample of about 5,000 persons located across the country, but 15 counties are surveyed annually.
Many will be familiar with some of the study's products such as standardized child growth charts (http://www.cdc.gov/growthcharts/cdc_charts.htm), or related initiatives such as the National Youth Fitness Survey (NNYFS) (http://www.cdc.gov/nchs/nnyfs.htm which collected data on physical activity and fitness levels in order to provide an evaluation of the health and fitness of children in the U.S. ages 3 to 15. 
The NNYFS collected data on physical activity and fitness levels of our youth through interviews and fitness tests.
Much of the NHANES and NNYFS data is available to the public through the National Center for Health Statistics at the U.S. Centers for Disease Control.
However, access to some data are restricted, available only to individuals via a Research Data Center (RDC) that can provide access.
<http://wwwn.cdc.gov/nchs/nhanes/search/variablelist.aspx?CycleBeginYear=1999&Component=Non-Public>.
There are 19 Federal Statistical Research Data Center (RDC) locations across the U.S.
A Google Scholar search shows more than 110,000 publications mention the NHANES dataset.
The Panel Study on Income Dynamics (PSID) \cite{PSID} is another large-scale population based effort.
PSID began in 1968 with a U.S. nationally representative sample of 5,000 families and more than 18,000 individuals.
The study consists of survey responses to questions about employment, income, wealth, expenditures, health, marriage, childbearing, child development, philanthropy, education, and other topics. As of mid-2015, PSID had generated more than 3,900 publications \cite{PSID_pubs}. 
Much of data is available publicly to users who register (http://simba.isr.umich.edu/U/ca.aspx) subject to the provisions of a formal set of use conditions (http://simba.isr.umich.edu/U/CondUse.aspx) that focus largely on ensuring that participants are not reidentified and that users properly cite and acknowledge use of the data. 
In addition, some restricted data elements are only available subject to a formal restricted use contract (http://simba.isr.umich.edu/restricted/RestrictedUse.aspx) between the investigator and the data owner, the University of Michigan.

In contrast the National Longitudinal Survey of Youth (NLSY) \cite{NLSY} has studied a large subset of the same cohort since the study began in 1979.
The project follows the lives of a sample of almost 10,000 American youth born between 1957-64.
Hosted by the U.S. National Bureau of Labor Statistics (BLS), NLSY ata are now available for up to 25 survey rounds on questions relating to employment and related questions.
The NLSY has also been extended to encompass women born in 1922-37 and 1943-55 and men born 1906-21 and 1941-52 through the National Longitudinal Survey (NLS).
Most NLSY and NLS data are available free of charge to by means of a web-based data portal \cite{NLSY_data}, access to which users must register for.
Users must apply for special permission to access geographic-related variables.
The Center for Human Resource Research (CHRR) at The Ohio State University and the National Opinion Research Center (NORC) at the University of Chicago administer both the data collection and repository functions.
A Google Scholar search in mid-2015 shows more than 20,000 publications cite the NLSY.

The Early Childhood Longitudinal Study (ECLS) is another large-scale population-based study that has followed the same groups of young people over time focusing on child development, school readiness, and early school experiences. 
The ECLS consists of three cohorts, one focused on a sample of children measured beginning at birth and two cohorts following children from kindergarten.
The birth cohort a sample of children born in 2001 and followed from birth through kindergarten entry, and unusual among comparable studies includes data from fathers. 
One of the kindergarten cohorts studied approximately 22,000 children who entered school in 1998-99 and were followe through the eighth grade. 
A second cohort entered kindergarten in 2010-11 and is being followed through the fifth grade.
The project has received funding from the National Center for Educational Statistics (https://nces.ed.gov/ecls/), and the data are archived at ICPSR.
ECLS consists of both public and restricted data, and according to a Google Scholar search has generated more than 8,000 publications, but ICSPR says 167.

Not all large-scale birth cohort studies succeed.
A notable example in the U.S. is the National Children's Study](http://www.nichd.nih.gov/research/NCS/Pages/default.aspx).
Authorized by the Children's Health Act of 2000, the NCS would have followed 100,000 children from before birth to age 21. 
However, the NIH Director decided to close the NCS on December 12, 2014 following the recommendations of an advisory panel.
Authorized by the Children’s Health Act of 2000, the NCS was a planned large-scale, long-term study of U.S. children and their mothers and fathers designed to study environmental influences on child health and development. 
Questionnaire, physical measures, biospecimens and environmental sample data from up to 5,726 participants in the NCS Vanguard study were collected in 2009-2014 prior to the NCS study's closure.
Those data are slated for release in a data archive sometime in 2015.

\subsubsection{Non-U.S. Examples}

Outside the U.S., there are several examples of large-scale longitudinal studies focused on children's growth and health.
Launched in 2011, the Japan Environment and Children's Study <http://www.env.go.jp/en/chemi/hs/jecs/> is a birth cohort study involving 100,000 parent-child pairs with the goal of evaluating the impact of various environmental factors on children's health and development. 
Pilot studies started in 2008 and the hundred thousandth participating mother was enrolled in March 2014. 
Health outcomes and exposure measurements will continue until the participating children become 13 years old.
The study is in its early stages, and has only produced a handful of publications <http://www.env.go.jp/en/chemi/hs/jecs/publications/index.html>.
Data sharing plans have not been announced.
World Health Organization (WHO) has offered to help coordinate measurement strategies for large-scale birth cohort studies <http://www.who.int/ceh/cohorts/en/> to ensure that data can be harmonized across them.

Several birth cohort studies have been undertaken in the U.K., beginning with the National Survey of Health & Development in 1946 (https://en.wikipedia.org/wiki/National_Survey_of_Health_%26_Development). 
The National Child Development Study (NCDS) is a continuing longitudinal study that seeks to follow the lives of all those living in Great Britain who were born in one particular week in 1958. 
The origins of the NCDS can be found in the Perinatal Mortality Survey (PMS) which was then sponsored by the National Birthday Trust Fund and set up to collect information about the social and obstetric factors associated with stillbirth and death in early infancy.[3] Examples of other topics which have since been included are medical care, health, home environment, educational progress, parental involvement, family relationships, economic activity, income, training and housing.[4]
Conducted by the Centre for Longitudinal Studies (CLS), the aim of the study is to improve understanding of the factors affecting human development over the whole lifespan. 
It collects information on physical and educational development, economic circumstances, employment, family life, health behaviour, wellbeing, social participation and attitudes.
physical, educational and social development.[1] During the period 2002-2004, genetic information on participants was also obtained to examine the genetic effects on common traits and diseases.[2]
The British Cohort Studies have generated more than 3,500 publications (http://www.cls.ioe.ac.uk/Bibliography.aspx?sitesectionid=647&sitesectiontitle=Bibliography) to-date.
Data are available from (http://ukdataservice.ac.uk/get-data/key-data/cohort-and-longitudinal-studies) by registration.
Another widely known and used dataset comes from the World Health Organization's (WHO) Multicentre Growth Reference Study (MGRS) (http://www.who.int/childgrowth/en/).
The MGRS collected growth and related data from 8500 children from widely different ethnic backgrounds and cultural settings, specifically Brazil, Ghana, India, Norway, Oman, and the U.S.
The data provide body mass index, height for children 5-19, and weight data norms for children 5-10.

The OECD-funded PISA dataset is available (http://pisa2000.acer.edu.au/index.php).
Data from the performance of students in Massive Open Online Courses at Harvard and MIT are available to the research community by application.

@@

Several themes emerge from this sample of population-based studies.
Datasets are large in terms of the volume of participants sampled ranging in the thousands to tens of thousands.
The datasets demonstrate significant variety of measures, including survey-based and bio-marker-based (NHANES, Ad Health, SECCYD) and even video (SECCYD).
The studies show variety geographic composition, as well.
Some studies involve tens of sample occasions per individual, the NLSY and Ad Health studies in particular.
Datasets are housed either in U.S. Government (NHANES, SECCYD) or university-based repositories (PSID, Ad Health, NLSY).
Most data are available to the public via web-based download or browsing/analysis portal, although access to data deemed sensitive may require a specific application and approval and possibly travel to a Research Data Center.
Finally, the datasets have generated research publications that range from the hundreds to the tens of thousands, suggesting that the collection, curation, and preservation of these resources has had a significant impact on scientific discovery.
% Susan McHale's comment about training. Check the other datasets for this.
% Limits on geographic linkage, other ways data are anomymized to reduce possibility of identifiability.

There are several large datasets with information about employment patterns and educational attainment in children and youth
The Bureau of Labor Statistics reports data about youth employment and unemployment patterns derived from the Census Bureau's Current Population Survey (http://www.bls.gov/cps/demographics.htm).
The U.S. Department of Education's National Center for Educational Statistics (NCES) collects and analyzes statistics on American primary, secondary, and post-secondary education, and many of the datasets are either public; data with individually identifiable data are available to researchers only under a restricted use agreement.

@@

@@

The next area of developmental science we explore is behavior genetics, where large data samples have long been required in order to detect what can often be small effect sizes.
Several large-scale genetics studies that are not specifically developmental deserve mention.
The Genome of the Netherlands Project(http://www.bbmri.nl/en-gb/activities/rainbow-projects/131-genome-of-the-netherlands) is a publicly available dataset from n=250 families, consisting of two parents plus one of their adult children \cite{Genome_Netherlands}.
The Genes, Environment, and Melanoma Study (GEMS) focuses on risk for melanoma in a sample of 3,700 individuals from the North America, Europe, and Australia (http://gemstudy.org/main/index.html).
The Psychiatric Genomics Consortium (PGC) (http://www.med.unc.edu/pgc) contains data from more than 170,000 individuals who have psychiatric diagnoses or are considered at-risk.
Many disorders featured (autism, attention-deficit hyperactivity disorder, and schizophrenia) in the PGC dataset have developmental dimensions.
Genome-wide association study (GWAS) results from the PGC dataset may be viewed in a specialized web browser tool (http://www.broadinstitute.org/mpg/ricopili/).

@@

The Twin and Offspring Study in Sweden (TOSS) \cite{TOSS} focuses on genetic and environmental contributions to measures of family relationships and mental health.
TOSS extends the Twin Moms Project to include a sample of twin fathers (N=320 twin pairs and their families).
In addition, TOSS plans to collect another 250 pairs of twin mothers for a target sample of 900 twin parents (n=3,000 individuals) and their families.
While still relatively early in data collection, more than 24 publications reference the TOSS dataset.
Data sharing plans have not yet been announced.

@@

Several large studies have focused on developmental patterns in brain structure or activity.
For practical reasons -- motion artifacts make imaging very young children extremely challenging -- most projects have focused on children and adolescents.
One of the earliest studies of its kind was a combined longitudinal and cross-sectional study of child and adolescent structural brain development led by Jay Giedd at NIMH \cite{giedd_brain_1999}.
The study collected structural brain information from 145 4 to 20-year-olds along with behavioral and clinical data.
More than 100 of the participants were scanned on more than one occasion, with the time between scans approximately two years.
The original paper describing this study has been cited more than 3,300 times, as of mid-2015.
The dataset has generated other highly cited papers \cite{gogtay_dynamic_2004}, but the data appear not to be available for analyses by researchers outside the original investigative team.
The NIH MRI Study of of Normal Brain Development \cite{NIH_Pediatric_MRI} collected multi-modal structural MRI from 554 children from 4 to 18 years of age.
A second cohort of children from birth to 4 years was scanned longitudinally with up to 10 scans. 
Demographic, hormonal, cognitive, affective, and psychiatric data were also collected.
The data are available to qualified researchers by application, and the dataset has generated some highly cited papers since data collection ended in 2006 \cite{group_total_2012, zielinski_network-level_2010}.
John Richards and colleagues have combined data from the NIH MRI Study, data collected in their own labs, and other public sources to create average structural brain templates that can provide more accurate bases for developmental functional neuroimaging studies using EEG and fMRI \cite{Richards_Developmental_MRI_Database, richards_database_2015}.
These data can be openly shared with appropriate citation.
More recently, the Pediatric Imaging, Neurocognition, and Genetics (PING) Project \cite{PING} collected multimodal neuroimaging data, genotypes, neurodevelopmental histories, and information about cognitive and social and emotional function in more than 1,000 participants 3-20 years of age recruited from Baltimore, Boston, Honolulu, Los Angeles, New Haven, New York, Sacramento, and San Diego. 
The study was funded by the Eunice Kennedy Shriver National Institute of Child Health and Human Development.
Data are available to the research community by application.
As of mid-2015, the project had generated more than 20 publications.
The Developmental Connectome Project \cite{developmental_connectome}, led by investigators at King’s College London, Imperial College London, and Oxford University goes against the trend of imaging studies that focus on older children.
This project aims to study the connectome, a map of human brain connectivity, from 20 to 44 weeks post-conception.
Funded by the European Research Council, the imaging data are accompanied by clinical, behavioral and genetic information.
Investigators hope that the data set will provide a basis for studying genetic and environmental risks that could lead to neurodevelopmental disorders such as Autistic Spectrum Disorder or Cerebral Palsy.
The project has already generated 17 publications as of mid-2015.

@@

There are several large-scale datasets associated with language processing.
The CHILDES/TalkBank \cite{macwhinney_childes_2001} archive is one of the largest and most well-established.
It consists of transcripts and audio and video recordings of children's utterances along with recent data from adults with aphasia.
CHILDES/TalkBank has generated more than ten thousand of citations, according to Google Scholar, and the data are available to the research community, some of them publicly.
The Human Speechome Project (http://www.media.mit.edu/cogmac/projects/hsp.html) at MIT's Media Lab involved recording 10 hours of audio and video from one child's home on a daily basis from birth to age three.
The project has generated more than 78 publications (http://www.media.mit.edu/cogmac/publications.html), but the data have not been made available outside of the original investigative team.
Data was gathered at an average rate of 200 gigabytes per day, necessitating the development of sophisticated data-mining tools to reduce analysis efforts to a manageable level, and transcribing significant speech added a labor-intensive dimension.
% edit previous
WordBank (http://wordbank.stanford.edu/) is an open access archive that consists of data from more than 40,000 adminstrations of the MacArthur-Bates Communicative Development Inventory (CDI).
Catherine Tamis-Lemonda has released video data (https://nyu.databrary.org/volume/8) from a longitudinal study of child language consisting of more than 1,000 sessions of infants and children (9 mos-7.6 years) and their mothers carrying out a series of semi-structured tasks.
The LENA Foundation has created a technology framework that allows children's speech in natural settings to be recorded and analyzed (http://www.lenafoundation.org/customer-resources/videos).
LENA stands for Language ENvironment Analysis.
The system provides an automatic language collection and analysis tool for speech language professionals and parents.
The system was designed and is distributed by the LENA Foundation.

>>>>

\subsection{Government-collected and Managed Datasets}

Some of the largest and most widely used datasets in developmental science are population-based, often with a focus on health.
The National Health and Nutrition Examination Survey (NHANES) \cite{NHANES} began collecting data in 1959 with a focus on characterizing the health and nutritional status of U.S. infants, children, and adults. 
The annual survey combines interviews about demographic, socioeconomic, dietary and health related topics and physical examinations involving medical, dental, and physiological measurements, as well as laboratory tests.
The study is cross-sectional and assesses a nationally representative sample of about 5,000 persons located across the country, with 15 counties surveyed annually.
Many will be familiar with some of NHANES's products such as standardized child growth charts (http://www.cdc.gov/growthcharts/cdc_charts.htm).
The National Youth Fitness Survey \cite{NYFS} is closely associated with NHANES.
NYFS collected data on physical activity and fitness levels in U.S. children ages 3 to 15 in 2012 in order to provide an evaluation their overall fitness.
Much of the NHANES and NNYFS data is available to the public through the National Center for Health Statistics at the U.S. Centers for Disease Control.
However, access to some data are restricted, available only to individuals via one of the 19 Federal Statistical Research Data Centers (RDC) locations across the U.S.
A Google Scholar search shows more than 110,000 publications mention the NHANES dataset.

The National Longitudinal Survey of Youth (NLSY) \cite{NLSY} provides an example of a large-scale study that has examined the same cohort -- almost 10,000 American youth born between 1957-64 -- on a annual basis since the study's beginning in 1979.
Hosted by the U.S. National Bureau of Labor Statistics (BLS), NLSY data are now available for up to 25 survey rounds on questions relating to employment and related issues.
The NLSY has also been extended to encompass women born in 1922-37 and 1943-55 and men born 1906-21 and 1941-52 through the National Longitudinal Survey (NLS).
Most NLSY and NLS data are available free of charge to by means of a web-based data portal \cite{NLSY_data} to which users must apply for access.
Access to geographic-related variables requires special permission.
The Center for Human Resource Research (CHRR) at The Ohio State University and the National Opinion Research Center (NORC) at the University of Chicago administer both the data collection and repository functions.
A Google Scholar search in mid-2015 shows that more than 20,000 publications cite the NLSY.

The Early Childhood Longitudinal Study (ECLS) is another large-scale population-based study that has followed the same groups of young people over time focusing on child development, school readiness, and early school experiences. 
The ECLS consists of three cohorts, one focused on a sample of children measured beginning at birth and two cohorts following children from kindergarten.
The birth cohort a sample of children born in 2001 and followed from birth through kindergarten entry, and unusual among comparable studies includes data from fathers. 
One of the kindergarten cohorts studied approximately 22,000 children who entered school in 1998-99 and were followe through the eighth grade. 
A second cohort entered kindergarten in 2010-11 and is being followed through the fifth grade.
The project has received funding from the National Center for Educational Statistics (https://nces.ed.gov/ecls/), and the data are archived at ICPSR.
ECLS consists of both publicly available and restricted data.
A Google Scholar search generates more than 8,000 hits, of which the ICPSR archive can verify 167 project-related publications (https://www.icpsr.umich.edu/icpsrweb/ICPSR/biblio/series/00191/resources?sortBy=1).

There are several large government-collected and shared datasets with information about employment patterns and educational attainment in children and youth.
The Bureau of Labor Statistics reports data about youth employment and unemployment patterns derived from the Census Bureau's Current Population Survey (http://www.bls.gov/cps/demographics.htm).
The U.S. Department of Education's National Center for Educational Statistics (NCES) collects and analyzes statistics on American primary, secondary, and post-secondary education, and many of the datasets are either public; data with individually identifiable data are available to researchers under a restricted use agreement.

There are several examples of large-scale longitudinal studies focused on children's growth, health, and education from outside the U.S.
The World Health Organization's (WHO) Multicentre Growth Reference Study (MGRS) \cite{MGRS} collected growth and related data from 8,500 children from widely different ethnic backgrounds and cultural settings, specifically Brazil, Ghana, India, Norway, Oman, and the U.S.
MGRS data provide body mass index, height for children 5-19, and weight data norms for children 5-10, in addition to other anthropomorphic measurements.
Launched in 2011, the Japan Environment and Children's Study \cite{JECS} is a birth cohort study involving 100,000 parent-child pairs with the goal of evaluating the impact of various environmental factors on children's health and development. 
Pilot studies started in 2008 and the hundred thousandth participating mother was enrolled in March 2014. 
Health outcomes and exposure measurements will continue until the participating children become 13 years old.
The study is in its early stages.
The study has produced a handful of publications \cite{JECS_pubs}, and no data sharing plans have been announced.
Several large-scale birth cohort studies have been undertaken in the U.K., including the National Survey of Health & Development \cite{NSHD_UK} and the National Child Development Study (NCDS) \cite{NCDS_UK}.
NCDS is an ongoing longitudinal study that seeks to follow the lives of all those living in Great Britain who were born in one particular week in 1958.
Originally focused on improving infant mortality, the study has expanded to focus on factors predicting wellbeing across the lifespan.
Measures include genetic, health, education, social, and economic factors.
The study is funded by the British government and data are available by registration from the U.K. Data Service \cite{UK_cohort_data}.
The British Cohort Studies have generated more than 3,500 publications to-date \cite{UK_cohort_citations}.
Finally, the Organization for Economic Cooperation and Development (OECD) publishes the PISA comparative educational dataset \cite{PISA}, which generates more than 5,000 search hits.

Not all large-scale birth cohort studies succeed.
A notable failure in the U.S. is the National Children's Study](http://www.nichd.nih.gov/research/NCS/Pages/default.aspx).
Authorized by the Children's Health Act of 2000, the NCS would have followed 100,000 children and their parents from before birth to age 21. 
However, the NIH Director decided to close the NCS on December 12, 2014 following the recommendations of an advisory panel. 
Questionnaire, physical measures, biospecimens and environmental sample data from up to 5,726 participants in the NCS Vanguard study were collected in 2009-2014 prior to the NCS study's closure.
Those data are slated for release in a data archive sometime in 2015.

@@

While no means exhaustive in its coverage, several themes emerge from this sample of U.S. and international population-based studies.
Datasets are large in terms of the volume of participants sampled, ranging in the thousands to tens of thousands.
Some studies involve tens of sample occasions per individual over extended periods of time, and show significant variety in the measures collected.
Most data are available to the public via web-based download or browsing/analysis portal, although access to data deemed sensitive may require a specific application and approval and possibly travel to a Research Data Center.
Finally, the datasets have generated research publications that range from the hundreds to the tens of thousands, suggesting that the collection, curation, and preservation of these resources has had a significant impact on scientific discovery.
We turn next to datasets whose collection and management is under the control of individual, usually academic investigators.

@@

The number and variety of topics studied

The Panel Study on Income Dynamics (PSID) \cite{PSID} is another large-scale population based effort.
PSID began in 1968 with a U.S. nationally representative sample of 5,000 families and more than 18,000 individuals.
The study consists of survey responses to questions about employment, income, wealth, expenditures, health, marriage, childbearing, child development, philanthropy, education, and other topics. As of mid-2015, PSID had generated more than 3,900 publications \cite{PSID_pubs}. 
Much of data is available publicly to users who register (http://simba.isr.umich.edu/U/ca.aspx) subject to the provisions of a formal set of use conditions (http://simba.isr.umich.edu/U/CondUse.aspx) that focus largely on ensuring that participants are not reidentified and that users properly cite and acknowledge use of the data. 
In addition, some restricted data elements are only available subject to a formal restricted use contract (http://simba.isr.umich.edu/restricted/RestrictedUse.aspx) between the investigator and the data owner, the University of Michigan.

@@

Large data samples have long been required in behavior genetics order to detect what can often be small effect sizes.
Several large-scale genetics studies that are not specifically developmental deserve mention.
The Genome of the Netherlands Project(http://www.bbmri.nl/en-gb/activities/rainbow-projects/131-genome-of-the-netherlands) is a publicly available dataset from n=250 families, consisting of two parents plus one of their adult children \cite{Genome_Netherlands}.
The Genes, Environment, and Melanoma Study (GEMS) focuses on risk for melanoma in a sample of 3,700 individuals from the North America, Europe, and Australia (http://gemstudy.org/main/index.html).
The Psychiatric Genomics Consortium (PGC) (http://www.med.unc.edu/pgc) contains data from more than 170,000 individuals who have psychiatric diagnoses or are considered at-risk.
Many disorders featured (autism, attention-deficit hyperactivity disorder, and schizophrenia) in the PGC dataset have developmental dimensions.
Genome-wide association study (GWAS) results from the PGC dataset may be viewed in a specialized web browser tool (http://www.broadinstitute.org/mpg/ricopili/).

@@

The Twin and Offspring Study in Sweden (TOSS) \cite{TOSS} focuses on genetic and environmental ontributions to measures of family relationships and mental health.
TOSS extends the Twin Moms Project to include a sample of twin fathers (N=320 twin pairs and their families).
In addition, TOSS plans to collect another 250 pairs of twin mothers for a target sample of 900 twin parents (n=3,000 individuals) and their families.
While still relatively early in data collection, more than 24 publications reference the TOSS dataset.
Data sharing plans have not yet been announced.

@@

Several large-scale studies have focused on developmental patterns in brain structure or activity.
For practical reasons -- motion artifacts make imaging very young children extremely challenging -- most projects have focused on children and adolescents.
One of the earliest studies of its kind was a combined longitudinal and cross-sectional study of child and adolescent structural brain development led by Jay Giedd at NIMH \cite{giedd_brain_1999}.
The study collected structural brain information from 145 4 to 20-year-olds along with behavioral and clinical data.
More than 100 of the participants were scanned on more than one occasion, with the time between scans approximately two years.
The original paper describing this study has been cited more than 3,300 times, as of mid-2015.
The dataset has generated other highly cited papers \cite{gogtay_dynamic_2004}, but the data appear not to be available for analyses by researchers outside the original investigative team.

The NIH MRI Study of of Normal Brain Development \cite{NIH_Pediatric_MRI} collected multi-modal structural MRI from 554 children from 4 to 18 years of age.
A second cohort of children from birth to 4 years was scanned longitudinally with up to 10 scans. 
Demographic, hormonal, cognitive, affective, and psychiatric data were also collected.
The data are available to qualified researchers by application, and the dataset has generated some highly cited papers since data collection ended in 2006 \cite{group_total_2012, zielinski_network-level_2010}.
John Richards and colleagues have combined data from the NIH MRI Study, data collected in their own labs, and other public sources to create average structural brain templates that can provide more accurate bases for developmental functional neuroimaging studies using EEG and fMRI \cite{Richards_Developmental_MRI_Database, richards_database_2015}.
These data can be openly shared with appropriate citation.

More recently, the Pediatric Imaging, Neurocognition, and Genetics (PING) Project \cite{PING} collected multimodal neuroimaging data, genotypes, neurodevelopmental histories, and information about cognitive and social and emotional function in more than 1,000 participants 3-20 years of age recruited from Baltimore, Boston, Honolulu, Los Angeles, New Haven, New York, Sacramento, and San Diego. 
The study was funded by the Eunice Kennedy Shriver National Institute of Child Health and Human Development.
Data are available to the research community by application.
As of mid-2015, the project had generated more than 20 publications.

The Developmental Connectome Project \cite{developmental_connectome}, led by investigators at King’s College London, Imperial College London, and Oxford University goes against the trend of imaging studies that focus on older children.
This project aims to study the connectome, a map of human brain connectivity, from 20 to 44 weeks post-conception.
Funded by the European Research Council, the imaging data are accompanied by clinical, behavioral and genetic information.
Investigators hope that the data set will provide a basis for studying genetic and environmental risks that could lead to neurodevelopmental disorders such as Autistic Spectrum Disorder or Cerebral Palsy.
The project has already generated 17 publications as of mid-2015.

@@

There are several large-scale datasets associated with language processing.
The CHILDES/TalkBank \cite{macwhinney_childes_2001} archive is one of the largest and most well-established.
It consists of transcripts and audio and video recordings of children's utterances along with recent data from adults with aphasia.
CHILDES/TalkBank has generated more than ten thousand of citations, according to Google Scholar, and the data are available to the research community, some of them publicly.

The Human Speechome Project (http://www.media.mit.edu/cogmac/projects/hsp.html) at MIT's Media Lab involved recording 10 hours of audio and video from one child's home on a daily basis from birth to age three.
The project has generated more than 78 publications (http://www.media.mit.edu/cogmac/publications.html), but the data have not been made available outside of the original investigative team.
Data was gathered at an average rate of 200 gigabytes per day, necessitating the development of sophisticated data-mining tools to reduce analysis efforts to a manageable level, and transcribing significant speech added a labor-intensive dimension.

% edit previous
WordBank (http://wordbank.stanford.edu/) is an open access archive that consists of data from more than 40,000 adminstrations of the MacArthur-Bates Communicative Development Inventory (CDI).
Catherine Tamis-Lemonda has released video data (https://nyu.databrary.org/volume/8) from a longitudinal study of child language consisting of more than 1,000 sessions of infants and children (9 mos-7.6 years) and their mothers carrying out a series of semi-structured tasks.

The LENA Foundation has created a technology framework that allows children's speech in natural settings to be recorded and analyzed (http://www.lenafoundation.org/customer-resources/videos).
LENA stands for Language ENvironment Analysis.
The system provides an automatic language collection and analysis tool for speech language professionals and parents.
The system was designed and is distributed by the LENA Foundation.
